### 2015-08-25 Proposals/NERC-CCUS_2015/Outline_ReactiveCCS.tex
Updates in the proposal
updates

\begin{enumerate}[(a)]
%
\item {\bf University of Aberdeen (UoA), School of Engineering - Environmental and Industrial Fluid Mechanics Group (EnIFluG)\footnote{\href{http://www.abdn.ac.uk/engineering/research/environmental-industrial-fluid-mechanics-122.php}{http://www.abdn.ac.uk/engineering/research/environmental-industrial-fluid-mechanics-122.php}}:} The Group specialises in combined computational methods, laboratory experimentation, and field work on permeable/porous media flows, sediment transport, coastal monitoring, turbulent buoyant jets and large-scale environmental flows. The Group has outstanding experimental facilities, including a range of large water. The latest addition is the petroleum engineering laboratory which represents a \pounds 200k investment in new equipment. The laboratory comprises four independent core-flood rigs with custom-made Hassler cells, an Anton Paar density meter and glass capillary viscometers with thermal bath. In May, UoA was awarded a further \pounds 950k (under the framework of Oil $\&$ Gas Academy of Scotland) for the purchase of equipment related to O$\&$G training. Among the equipment that will be purchased is an X-ray micro-computed tomography (microCT) system for in-situ imaging of fluids in rock samples. \red{make it more general}. \\
{\bf Dr Jefferson Gomes (JG)} leads the development of the computational multi-physics models within Fluidity. He published >25 journal papers on CFD, porous media flows, fluid phase equilibria, nuclear criticality and computational optimisation.

@@

%%%
%%% SUBSECTION
%%%
\subsection{Motivation and Objectives}
\CO\; has been injected in underground geological formations to enhance hydrocarbon productivity (e.g. \CO-EOR). %whilst disposing acidic gasses (mainly H$_{2}$S and \CO) from natural occurance or captured from industrial plants. 
This has helped to prove that geological \CO\; storage is a viable engineering and economic strategy for reducing GHG emissions. However, geological \CO\; storage schemes must comply with UK/EUrequirements (European Council's directive 2009/31/EC) for:
\begin{enumerate}[(a)]
%
   \item Site screening and selection to identify potential geological locations for \CO\; storage;\label{Motivation:a}
%
   \item Site characterisation to assess the engineering and economic feasibility;\label{Motivation:b}
%
   \item Simulation of potential injection strategies to assess (i) pressure and stress distribution and their evolution in time, (ii) plume migration, (iii) oil recovery, (iv) storage capacity and (v) flow rate of injection;\label{Motivation:c}
%
   \item Risk assessment to identify risks of leakage via pathways identified in (\ref{Motivation:c}), to design a monitoring scheme for continuous assessment of these risks, and to develop contingency plans in case of  leakage scenarios;\label{Motivation:d}
%
   \item Site management and operation to ensure that the site is operated in compliance with regulations and to continuously assess the risks associated with the storage site during operation, closure and post-closure stages.\label{Motivation:e} 
%
\end{enumerate}

@@

The main objectives of this proposal are:
\begin{enumerate}[(i)]
%
   \item Investigation of \CO\; plume dynamics (migration and leakage);
%
   \item \red{Investigation of multi-fluid flows through zones of potential rheological weakness in the overburden (e.g., faults and chimneys): solid-fluid interactions and stress-induced fractures formation;}
%
   \item \red{Model uncertainty quantification: sensor network (data assimilation) and ROM};
%
   \item Model quality assurance (monitoring $\&$ verification for risk assessment and environmental impact).
%
\end{enumerate}

{\bf The novelty of this project is in addressing in an integrated way detailed modelling of \CO\; flows under reservoir conditions (plume migration $\&$ leakage), changes on petrophysical properties (permeability/porosity spatial- and time-variability), solid-fluid interactions (integrity of repositories) and monitoring $\&$ verification technologies.} We shall consider multi-scale and multi-physics aspects of CCUS problem: two length-scales:

>>>>

\begin{enumerate}[(a)]
%
\item {\bf University of Aberdeen (UoA), School of Engineering - Environmental and Industrial Fluid Mechanics Group (EnIFluG)\footnote{\href{http://www.abdn.ac.uk/engineering/research/environmental-industrial-fluid-mechanics-122.php}{http://www.abdn.ac.uk/engineering/research/environmental-industrial-fluid-mechanics-122.php}}:} The Group specialises in combined computational methods, laboratory experimentation, and field work on permeable/porous media flows, sediment transport, coastal monitoring, turbulent buoyant jets and large-scale environmental flows. The Group has outstanding experimental facilities, including a range of large water. The latest addition is the petroleum engineering laboratory which represents a \pounds 200k investment in new equipment. The laboratory comprises four independent core-flood rigs with custom-made Hassler cells, an Anton Paar density meter and glass capillary viscometers with thermal bath. In May, UoA was awarded a further \pounds 950k (under the framework of Oil $\&$ Gas Academy of Scotland) for the purchase of equipment related to O$\&$G training. Among the equipment that will be purchased is an X-ray micro-computed tomography (microCT) system for in-situ imaging of fluids in rock samples. \red{make it more general}. \\
{\bf Dr Jefferson Gomes (JG)} leads the development of the computational multi-physics models within Fluidity. He published >25 journal papers on CFD, porous media flows, fluid phase equilibria, nuclear criticality and computational optimisation.\\
{\bf Dr Yukie Tanino (YT)} research combines theory and laboratory experiment to model flow and mass transport in obstructed and/or confined systems, including geological reservoirs, inclined pipes, and artificial wetlands for wastewater treatment. She manages the recently established petroleum engineering laboratory in the School.

@@

%%%
%%% SUBSECTION
%%%
\subsection{Motivation and Objectives}
\CO\; has been injected in underground geological formations to enhance hydrocarbon productivity (e.g. \CO-EOR). %whilst disposing acidic gasses (mainly H$_{2}$S and \CO) from natural occurance or captured from industrial plants. 
This has helped to prove that geological \CO\; storage is a viable engineering and economic strategy for reducing GHG emissions. However, geological \CO\; storage schemes must comply with UK/EUrequirements (European Council's directive 2009/31/EC) for:
\begin{enumerate}[(i)]
%
   \item Site screening and selection to identify potential geological locations for \CO\; storage;\label{Motivation:a}
%
   \item Site characterisation to assess the engineering and economic feasibility;\label{Motivation:b}
%
   \item Simulation of potential injection strategies to assess (i) pressure and stress distribution and their evolution in time, (ii) plume migration, (iii) oil recovery, (iv) storage capacity and (v) flow rate of injection;\label{Motivation:c}
%
   \item Risk assessment to identify risks of leakage via pathways identified in (\ref{Motivation:c}), to design a monitoring scheme for continuous assessment of these risks, and to develop contingency plans in case of  leakage scenarios;\label{Motivation:d}
%
   \item Site management and operation to ensure that the site is operated in compliance with regulations and to continuously assess the risks associated with the storage site during operation, closure and post-closure stages.\label{Motivation:e} 
%
\end{enumerate}

@@

The main objectives of this proposal are: 
\begin{enumerate}[(a)]
%
   \item Investigation of \CO\; plume dynamics (migration and leakage);
%
   \item \red{Investigation of multi-fluid flows through zones of potential rheological weakness in the overburden (e.g., faults and chimneys): solid-fluid interactions and stress-induced fractures formation;}
%
   \item \red{Model uncertainty quantification: sensor network (data assimilation) and ROM};
%
   \item Model quality assurance (monitoring $\&$ verification for risk assessment and environmental impact).
%
\end{enumerate}

{\bf The novelty of this project is in addressing in an integrated way: detailed modelling of \CO\; flows under reservoir conditions (plume migration $\&$ leakage); changes on petrophysical properties (permeability/porosity spatial- and time-variability); solid-fluid interactions (integrity of repositories); and monitoring $\&$ verification technologies.}

\medskip
We shall consider two integrated length-scales in the CCUS problem: 
\begin{enumerate}[(a)]
  \item \red{flow mechanisms and structures in pore-scale} -- {\bf YT};
  \item flow detachment in field scale \blue{(develop this as a backbone for phase change and flow instability)} -- {\bf JG}.
\end{enumerate}
